---
title: "Data preperation and analysis of part III of basal attributes article"
author: "Julius Fenn"
format:
  html:
    toc: true
    toc-depth: 3
    html-math-method: katex
    number-sections: true
---


# Notes


# global variables

Define your global variables (can take some time to run):

```{r}
## global variables
runMplusLCA <- FALSE
## set to lower values to reduce compilation time: 
nRep_def = 100 # 100
n.lambda_def = 40 # 40
LCArunsDef = 6 # 10
```


# create raw data files



```{r}
#| label: create raw files
#| warning: false

# sets the directory of location of this script as the current directory
# setwd(dirname(rstudioapi::getSourceEditorContext()$path))

### load packages
require(pacman)
p_load('tidyverse', 'jsonlite',
       'stargazer',  'DT', 'psych',
       'writexl',
       'lavaan', 'MplusAutomation', 'semPlot',
       'ggstatsplot')


### load socio-demographic data
setwd("data demographic")
prolific <- read.csv(file = "prolific_export_6830619c7e8edcf34faa12ff.csv", header = TRUE)

### load JATOS data
setwd("../data")
suppressMessages(
  read_file('jatos_results_data_20250602071616.txt') %>%
    # ... split it into lines ...
    str_split('\n') %>% first() %>%
    # ... filter empty rows ...
    discard(function(x)
      x == '') %>%
    discard(function(x)
      x == '\r') %>%
    # ... parse JSON into a data.frame
    map_dfr(fromJSON, flatten = TRUE)
) -> dat

#> add ID counter
dat$ID <- NA

tmp_IDcounter <- 0
for (i in 1:nrow(dat)) {
  if (!is.na(dat$sender[i]) &&
      dat$sender[i] == "Greetings") {
    tmp_IDcounter = tmp_IDcounter + 1
  }
  dat$ID[i] <- tmp_IDcounter
}
rm(tmp_IDcounter)


### load functions
setwd("../functions")
for(i in 1:length(dir())){
  # print(dir()[i])
  source(dir()[i], encoding = "utf-8")
}

rm(i)
```


# set up and save data

## get overview over data


```{r}
### keep only complete data sets
sort(table(dat$ID))
sum(table(dat$ID) != max(table(dat$ID)))
sum(table(dat$ID) == max(table(dat$ID)))
```


```{r}
dim(dat)
dat <-
  dat[dat$ID %in% names(table(dat$ID))[table(dat$ID) == max(table(dat$ID))], ]
dim(dat)
```

```{r}
str_subset(string = colnames(dat), pattern = "^ATI|^HEXACO")
str_subset(string = colnames(dat), pattern = "^HEXACO")
```
## create data files

Set up questionnaire:

```{r}
#| label: create questionnaires files
#| warning: false


### json (from JATOS) to 2D data.frame
# > pre study
# add paradata
tmp_notNumeric <- c(str_subset(string = colnames(dat), pattern = "^meta"))
tmp_notNumeric <-
  str_subset(string = tmp_notNumeric,
             pattern = "labjs|location",
             negate = TRUE)

### get survey
vec_ques <- c("PROLIFIC_PID",
              "dummy_informedconsent",
              "commCheck",
              str_subset(string = colnames(dat), pattern = "^ATI|^HEXACO"),
              "feedback_conscientiousCompletion",
              tmp_notNumeric)

vec_notNumeric = c("PROLIFIC_PID", "feedback_critic", tmp_notNumeric)

questionnaire <- questionnairetype(
  dataset = dat,
  listvars = vec_ques,
  notNumeric = vec_notNumeric,
  verbose = FALSE
)

dim(questionnaire)
```
Add Prolific data to data set:

```{r}
### add prolific data
prolific <- prolific[prolific$Participant.id %in% questionnaire$PROLIFIC_PID,]
prolific <- prolific %>%
  arrange(sapply(Participant.id, function(y) which(y == questionnaire$PROLIFIC_PID)))



if(all(prolific$Participant.id == questionnaire$PROLIFIC_PID)){
  print("prolific data sucessfully added")

  questionnaire$socio_age <- prolific$Age
  questionnaire$socio_sex <- prolific$Sex
  questionnaire$socio_ethnicity <- prolific$Ethnicity.simplified
  questionnaire$socio_student <- prolific$Student.status
  questionnaire$socio_employment <- prolific$Employment.status
  questionnaire$total_min_prolific <- prolific$Time.taken / 60

  questionnaire$socio_age <- as.numeric(questionnaire$socio_age)
}
```
Get Ratings of Vignettes:

```{r}
# Pre-define the column names with rat_ pattern
rat_cols <- str_subset(string = colnames(dat), pattern = "^rat_")

# Create an empty list to store processed data
result_list <- list()

# Loop over each unique ID
for (i in unique(dat$ID)) {
  tmp <- dat[dat$ID == i, c("ID", "PROLIFIC_PID", "Vignette", rat_cols)]
  
  # Fill down PROLIFIC_PID if missing
  tmp <- tmp %>%
    fill(PROLIFIC_PID, .direction = "downup")
  
  # Filter rows where all rat_ columns are not NA
  tmp <- tmp %>%
    filter(if_any(all_of(rat_cols), ~ !is.na(.)))
  
  # Append to result list
  result_list[[as.character(i)]] <- tmp
}

# Combine all into one dataframe
questionnaire_rating_long <- bind_rows(result_list)
questionnaire_rating_long$Vignette <- factor(questionnaire_rating_long$Vignette)

questionnaire_rating_long <- questionnaire_rating_long %>%
  mutate(across(starts_with("rat_"), as.numeric))

questionnaire_rating_long$rat_mean <- rowMeans(x = questionnaire_rating_long[, str_subset(string = colnames(questionnaire_rating_long), pattern = "^rat_")])





# Assuming your dataset is named questionnaire_rating_long
questionnaire_rating_wide <- questionnaire_rating_long %>%
  pivot_wider(
    id_cols = c(ID, PROLIFIC_PID),
    names_from = Vignette,
    values_from = c(rat_helpful, rat_trust, rat_buy, rat_emotion, rat_mean),
    names_glue = "{.value}_v{Vignette}"
  )

```


# analyze data

## describe sample

```{r}
table(questionnaire$socio_sex)
round(x = table(questionnaire$socio_sex) / nrow(questionnaire) * 100, digits = 2)

psych::describe(x = questionnaire[, c("socio_age", "total_min_prolific")])
```


## analyze and summarize surveyed scales

```{r}
psych::cor.plot(r = cor(questionnaire[, str_subset(string = colnames(questionnaire), pattern = "^ATI")]))
```
```{r}
questionnaire$`ATI-6` <- 7 - questionnaire$`ATI-6`
questionnaire$`ATI-3` <- 7 - questionnaire$`ATI-3`
questionnaire$`ATI-8` <- 7 - questionnaire$`ATI-8`
```


```{r}
psych::cor.plot(r = cor(questionnaire[, str_subset(string = colnames(questionnaire), pattern = "^ATI")]))
```
```{r}
questionnaire$mean_ATI <- rowMeans(questionnaire[, str_subset(string = colnames(questionnaire), pattern = "^ATI")])


questionnaire_rating_wide$mean_ATI <- questionnaire$mean_ATI
```








## analyze and summarize ratings

### overall EFA

```{r}
# ...
```



### EFA, CFA single scales

```{r}
setwd("outputs/CFAs")
#### Vignette 1
out_V1 <- CFAstats(dataset = questionnaire_rating_wide ,
                           regularExp = "^(?!rat_mean).*_v1$",
                           labelLatent = "Vignette1",
                           showPlots = TRUE, computeEFA = TRUE, computeCFA = TRUE, computeCFAMplus = TRUE)

#### Vignette 2
out_V2 <- CFAstats(dataset = questionnaire_rating_wide ,
                           regularExp = "^(?!rat_mean).*_v2$",
                           labelLatent = "Vignette2",
                           showPlots = TRUE, computeEFA = TRUE, computeCFA = TRUE, computeCFAMplus = TRUE)

#### Vignette 3
out_V3 <- CFAstats(dataset = questionnaire_rating_wide ,
                           regularExp = "^(?!rat_mean).*_v3$",
                           labelLatent = "Vignette3",
                           showPlots = TRUE, computeEFA = TRUE, computeCFA = TRUE, computeCFAMplus = TRUE)

#### Vignette 4
out_V4 <- CFAstats(dataset = questionnaire_rating_wide ,
                           regularExp = "^(?!rat_mean).*_v4$",
                           labelLatent = "Vignette4",
                           showPlots = TRUE, computeEFA = TRUE, computeCFA = TRUE, computeCFAMplus = TRUE)

#### Vignette 5
out_V5 <- CFAstats(dataset = questionnaire_rating_wide ,
                           regularExp = "^(?!rat_mean).*_v5$",
                           labelLatent = "Vignette5",
                           showPlots = TRUE, computeEFA = TRUE, computeCFA = TRUE, computeCFAMplus = TRUE)

#### Vignette 6
out_V6 <- CFAstats(dataset = questionnaire_rating_wide ,
                           regularExp = "^(?!rat_mean).*_v6$",
                           labelLatent = "Vignette6",
                           showPlots = TRUE, computeEFA = TRUE, computeCFA = TRUE, computeCFAMplus = TRUE)
```




### CFA first order

```{r}
model <- '
  # Measurement model
  Vignette1 =~ rat_helpful_v1 + rat_trust_v1 + rat_buy_v1 + rat_emotion_v1
  Vignette2 =~ rat_helpful_v2 + rat_trust_v2 + rat_buy_v2 + rat_emotion_v2
  Vignette3 =~ rat_helpful_v3 + rat_trust_v3 + rat_buy_v3 + rat_emotion_v3
  Vignette4 =~ rat_helpful_v4 + rat_trust_v4 + rat_buy_v4 + rat_emotion_v4
  Vignette5 =~ rat_helpful_v5 + rat_trust_v5 + rat_buy_v5 + rat_emotion_v5
  Vignette6 =~ rat_helpful_v6 + rat_trust_v6 + rat_buy_v6 + rat_emotion_v6
'



# Fit the SEM model
fit <- sem(model, data = questionnaire_rating_wide, meanstructure = FALSE, std.lv = TRUE)

# Print summary with standardized estimates
summary(fit, standardized = TRUE, fit.measures = TRUE)

head(modificationindices(fit, sort=T))


# Plot the SEM model
semPaths(fit,
         what = "std",            # show standardized estimates
         layout = "tree",         # neat tree layout
         edge.label.cex = 0.8,    # size of edge labels
         sizeMan = 5,             # manifest variable size
         sizeLat = 6,             # latent variable size
         fade = FALSE,            # keep all paths visible
         nCharNodes = 0,          # do not truncate labels
         rotation = 1)            # vertical tree layout


# Extract standardized latent variable correlations
latent_cor <- lavInspect(fit, "cor.lv")
print(latent_cor)

corPlot(latent_cor,
        numbers = TRUE,
        main = "Latent Correlations",
        zlim = c(-1, 1),
        cex = 1)

cor.plot(r = (cor(questionnaire_rating_wide[,sort(str_subset(string = colnames(questionnaire_rating_wide), pattern = "^rat_mean"))])),
                 numbers = TRUE,
        main = "Manifest Correlations",
        zlim = c(-1, 1),
        cex = 1)
```







## latent class analysis for "B: Grid (Hierarchy-Egalitarianism)"

over complete data set:

```{r}
setwd("outputs/LCA")


if(runMplusLCA){
  

  LCA_dat <- questionnaire_rating_wide[, c("ID",
                                      str_subset(string = colnames(questionnaire_rating_wide), pattern = "^rat_mean"))]
  tmp <- str_subset(string = colnames(LCA_dat), pattern = "^rat_mean_")
  tmp <- str_remove_all(string = tmp, pattern = "^rat_mean_")
  # cat(sort(tmp))
  # cat(paste0(sort(tmp), "(", 1:length(tmp), ")"))

  colnames(LCA_dat) <- c("ID", tmp)


  # prepareMplusData(df = LCA_dat, filename = "LCA_dat.dat")


  l = 1
  list_lca <- list()
  for(i in 2:LCArunsDef){
    print(i)

    numClasses <- i

    LCA_mplus  <- mplusObject(

      TITLE = paste0("Latent Class Analysis Mean Rating Vignettes", " c=", numClasses),

      VARIABLE =paste0("
  usevariables = v1 v2 v3 v4 v5 v6;

  classes      = c(", numClasses, ")"),

  ANALYSIS =
    "
    Type=mixture; ! LCA analysis
    STARTS= 500 100;
    !LRTstarts=0 0 300 20;
  ",

  PLOT =
    "
    type = plot3;
    series is v1(1) v2(2) v3(3) v4(4) v5(5) v6(6);
  ",

  SAVEDATA = paste0("file = lca_", numClasses, ".txt ;
    save = cprob;
    format = free;
  "),

  OUTPUT = "tech11 tech14;", rdata = LCA_dat)

    list_lca[[l]] <- mplusModeler(LCA_mplus,
                                         modelout = paste0("list_lca", numClasses, ".inp"),
                                         run = 1L)

    l = l + 1
  }

  # comment out to avoid overwriting LCA outputs
  saveRDS(list_lca, file="list_lca.rds")

}else{
  list_lca <- readRDS("list_lca.rds" )
}
```

get fit statistics LCA:

```{r}
getLCAfitstatistics(listMplusOutput = list_lca)
```


get profile plot:

```{r}
### number of classes
num_classes <- 3 # index - 1

# Step 1: Extract means
params <- list_lca[[num_classes]]$results$parameters$unstandardized

means_df <- params %>%
  filter(paramHeader == "Means" & !grepl("^C#\\d$", param))

# Step 2: Get class proportions
class_counts <- table(list_lca[[num_classes]]$results$savedata$C)
total_n <- sum(class_counts)
class_props <- round(100 * class_counts / total_n, 1)

# Create descriptive labels
class_labels <- paste0("Class ", names(class_counts), " (", class_props, "%)")

# Map class numbers to labels
class_label_map <- setNames(class_labels, paste0("Class ", names(class_counts)))

# Step 3: Prepare data for plotting
means_long <- means_df %>%
  mutate(Indicator = param,
         Class = paste0("Class ", LatentClass)) %>%
  select(Indicator, Class, est) %>%
  pivot_wider(names_from = Class, values_from = est) %>%
  pivot_longer(cols = starts_with("Class"), names_to = "Class", values_to = "Mean")

# Apply the new labels
means_long$Class <- class_label_map[means_long$Class]

# Optional: Set indicator order
# means_long$Indicator <- factor(means_long$Indicator,
#                                 levels = paste0("Vignette", 1:6))

# Step 4: Plot
ggplot(means_long, aes(x = Indicator, y = Mean, group = Class, color = Class)) +
  geom_line(size = 1) +
  geom_point(size = 2) +
  theme_minimal() +
  labs(title = "Latent Profiles",
       x = "Vignettes",
       y = "Estimated Means") +
  theme(text = element_text(size = 14),
        axis.text.x = element_text(angle = 45, hjust = 1),
        legend.title = element_blank())


class_counts
```






```{r}
questionnaire_rating_wide$classAssignment <- list_lca[[num_classes]]$results$savedata$C
boxplot(questionnaire_rating_wide$rat_mean_v1 ~ questionnaire_rating_wide$classAssignment)
boxplot(questionnaire_rating_wide$rat_mean_v6 ~ questionnaire_rating_wide$classAssignment)
```
```{r}
boxplot(questionnaire_rating_wide$mean_ATI ~ questionnaire_rating_wide$classAssignment)
summary(aov(questionnaire_rating_wide$mean_ATI ~ questionnaire_rating_wide$classAssignment))


summary_stats <- questionnaire_rating_wide %>%
  group_by(classAssignment) %>%
  summarise(
    N = n(),
    across(
      .cols = starts_with("mean_ATI"),
      .fns = list(mean = ~ mean(as.numeric(.), na.rm = TRUE),
                  sd = ~ sd(as.numeric(.), na.rm = TRUE)),
      .names = "{.col}_{.fn}"
    )
  )

summary_stats
```


```{r}
questionnaire_rating_long$classAssignment <- NA


for(i in 1:length(unique(questionnaire_rating_long$PROLIFIC_PID))){
  questionnaire_rating_long$classAssignment[questionnaire_rating_long$PROLIFIC_PID %in% unique(questionnaire_rating_long$PROLIFIC_PID)[i]] <- list_lca[[num_classes]]$results$savedata$C[i]
}
```


```{r}
summary_stats <- questionnaire_rating_long %>%
  group_by(Vignette, classAssignment) %>%
  summarise(
    N = n(),
    across(
      .cols = starts_with("rat_mean"),
      .fns = list(mean = ~ mean(as.numeric(.), na.rm = FALSE),
                  sd = ~ sd(as.numeric(.), na.rm = FALSE)),
      .names = "{.col}_{.fn}"
    )
  )

summary_stats
```



```{r}
# Pivot longer for ggplot
summary_long <- summary_stats %>%
  pivot_longer(
    cols = -c(Vignette, N),
    names_to = c("Rating", ".value"),
    names_pattern = "(rat_.*)_(mean|sd)"
  )

ggplot(summary_long, aes(x = Vignette, y = mean, fill = Rating)) +
  geom_col(position = position_dodge(width = 0.9), width = 0.7) +
  geom_errorbar(
    aes(ymin = mean - sd, ymax = mean + sd),
    position = position_dodge(width = 0.9),
    width = 0.2
  ) +
  facet_wrap(~ Rating, scales = "free_y") +
  labs(
    title = "Mean Ratings per Vignette",
    y = "Mean Rating ± SD",
    x = "Vignette"
  ) +
  theme_minimal() +
  theme(legend.position = "none")

```

```{r}
# library(ggstatsplot)
# 
# One plot for a single rating variable (within-subjects design)
# ggwithinstats(
#   data = questionnaire_rating_long,
#   x = Vignette,
#   y = rat_mean,
#   id = PROLIFIC_PID,  # Participant ID for within-subjects
#   type = "parametric",  # or "nonparametric" based on assumption checks
#   title = "Within-Subjects Comparison of Mean Ratings Across Vignettes",
#   xlab = "Vignette",
#   ylab = "Mean Ratings",
#   messages = FALSE
# )
```

```{r}
library(afex)
# Run repeated measures ANOVA (within-subjects)
anova_result <- aov_ez(
  id = "PROLIFIC_PID",
  dv = "rat_mean",
  data = questionnaire_rating_long,
  within = "Vignette",   # <- not between
  type = 3
)

summary(anova_result)


library(emmeans)

# Get estimated marginal means and post hoc comparisons
emmeans_result <- emmeans(anova_result, pairwise ~ Vignette)

emmeans_result$contrasts

```


```{r}
setwd("outputs")

library(lme4)
library(performance)

# Fit base model
model_base <- lmer(
  rat_mean ~ 1 +
    (1 | PROLIFIC_PID),
  data = questionnaire_rating_long
)

# ICC calculation
var_comp <- as.data.frame(VarCorr(model_base))
var_participant <- var_comp$vcov[var_comp$grp == "PROLIFIC_PID"]
var_residual <- var_comp$vcov[var_comp$grp == "Residual"]
icc <- var_participant / (var_participant + var_residual)
print(icc)

# Make sure Vignette is a factor
questionnaire_rating_long$Vignette_ref <- factor(questionnaire_rating_long$Vignette)

# Set level 6 as reference
questionnaire_rating_long$Vignette_ref <- relevel(questionnaire_rating_long$Vignette_ref, ref = "6")

# Fit the model with Vignette as fixed effect and random intercepts per participant
model_vignette <- lmer(
  rat_mean ~ Vignette_ref +
    (1 | PROLIFIC_PID),
  data = questionnaire_rating_long
)

# Use sjPlot instead of stargazer for mixed models
library(sjPlot)
tab_model(model_vignette, file = "lmm_vignette.html")


# Show summary
summary(model_vignette)

# Calculate R-squared
r2_values <- model_performance(model_vignette)
print(r2_values)
```



## try dummies:









```{r}
questionnaire_rating_long$dummyPart1 <- 0
questionnaire_rating_long$dummyPart2 <- 0
questionnaire_rating_long$dummyPart3 <- 0
questionnaire_rating_long$dummyPart4 <- 0
questionnaire_rating_long$dummyPart5 <- 0
questionnaire_rating_long$dummyPart6 <- 0


questionnaire_rating_long$dummyPart1[questionnaire_rating_long$Vignette %in% c(3,4,5)] <- 1
questionnaire_rating_long$dummyPart2[questionnaire_rating_long$Vignette %in% c(5,6)] <- 1
questionnaire_rating_long$dummyPart3[questionnaire_rating_long$Vignette %in% c(1,2)] <- 1
questionnaire_rating_long$dummyPart4[questionnaire_rating_long$Vignette %in% c(6)] <- 1
questionnaire_rating_long$dummyPart5[questionnaire_rating_long$Vignette %in% c(2,4)] <- 1
questionnaire_rating_long$dummyPart6[questionnaire_rating_long$Vignette %in% c(1,2)] <- 1

```


```{r}
summary_stats <- questionnaire_rating_long %>%
  group_by(dummyPart1, dummyPart2, dummyPart3, dummyPart4, dummyPart5, dummyPart6) %>%
  summarise(
    N = n(),
    across(
      .cols = starts_with("rat_mean"),
      .fns = list(mean = ~ mean(as.numeric(.), na.rm = TRUE),
                  sd = ~ sd(as.numeric(.), na.rm = TRUE)),
      .names = "{.col}_{.fn}"
    )
  )

summary_stats


```



```{r}
dummy_matrix <- questionnaire_rating_long[, c("dummyPart1", "dummyPart2", "dummyPart3",
                                              "dummyPart4", "dummyPart5", "dummyPart6")]
qr(dummy_matrix)$rank   # Actual rank
ncol(dummy_matrix)      # Number of columns

library(caret)
findLinearCombos(dummy_matrix)

```



```{r}
library(lme4)
model_base <- lmer(
  rat_mean ~ dummyPart1 + dummyPart2 + dummyPart3 + dummyPart4 + dummyPart5 + dummyPart6 +
    (1 | PROLIFIC_PID),
  data = questionnaire_rating_long
)

summary(model_base)
```




```{r}
library(psych)
library(dplyr)
library(ggplot2)
library(cluster)
library(factoextra)  # for fviz_cluster, optional visualization

rat_vars <- c("rat_helpful", "rat_trust", "rat_buy", "rat_emotion")
unique_vignettes <- sort(unique(questionnaire_rating_long$Vignette))

for (v in unique_vignettes) {
  cat("\n=== Vignette:", v, "===\n")
  
  # Subset and clean data
  vignette_data <- questionnaire_rating_long %>%
    filter(Vignette == v) %>%
    select(all_of(rat_vars)) %>%
    na.omit()
  
  # Check if enough data
  if (nrow(vignette_data) >= 5) {
    
    # 1. Correlation matrix plot
    cat("Correlation matrix:\n")
    corPlot(cor(vignette_data), main = paste("Correlation Plot - Vignette", v))
    
    # 2. Parallel analysis (scree plot)
    cat("Parallel analysis:\n")
    efa_parallel <- fa.parallel(vignette_data, fa = "fa", fm = "ml", main = paste("Parallel Analysis - Vignette", v))
    
    # 3. Optional: Run EFA with 1 factor if unidimensional
    efa <- fa(vignette_data, nfactors = efa_parallel$nfact, fm = "ml")
    print(efa$loadings)
    
     # 4. K-means clustering (set k = 2 or pick adaptively)
    vignette_scaled <- scale(vignette_data)

    cat("Selecting optimal number of clusters (k):\n")
    # Use elbow method to find optimal k
    nb <- fviz_nbclust(
      vignette_scaled,
      FUNcluster = kmeans,
      method = "wss",
      k.max = 6,
      nstart = 25  # needed by kmeans
    )
    
    print(nb)
    
    # Extract optimal k (visually, or use function if needed)
    optimal_k <- which.min(diff(diff(nb$data$y))) + 1
    cat("Auto-selected k =", optimal_k, "\n")
    
    set.seed(123)
    km <- kmeans(vignette_scaled, centers = optimal_k, nstart = 25)
    
    cat("K-means clustering results (k =", optimal_k, "):\n")
    print(km$size)
    print(km$centers)
    
    # 5. Plot clusters using PCA
    fviz_cluster(km, data = vignette_scaled, main = paste("K-Means Clusters - Vignette", v))
    
    
    
  } else {
    cat("Not enough complete responses for Vignette", v, "\n")
  }
}

```












```{r}
library(psych)
library(dplyr)
library(ggplot2)
library(cluster)
library(factoextra)

rat_vars <- c("rat_helpful", "rat_trust", "rat_buy", "rat_emotion")
unique_vignettes <- sort(unique(questionnaire_rating_long$Vignette))

# === NEW: list to store assignments ===
cluster_assignments <- list()

for (v in unique_vignettes) {
  cat("\n=== Vignette:", v, "===\n")
  
  # Subset and clean data
  vignette_subset <- questionnaire_rating_long %>%
    filter(Vignette == v) %>%
    select(PROLIFIC_PID, all_of(rat_vars)) %>%
    na.omit()
  
  # Keep PID separate and scale only the rating data
  pids <- vignette_subset$PROLIFIC_PID
  vignette_data <- vignette_subset %>% select(-PROLIFIC_PID)
  
  if (nrow(vignette_data) >= 5) {
    
    # 1. Correlation matrix plot
    cat("Correlation matrix:\n")
    corPlot(cor(vignette_data), main = paste("Correlation Plot - Vignette", v))
    
    # 2. Parallel analysis
    cat("Parallel analysis:\n")
    efa_parallel <- fa.parallel(vignette_data, fa = "fa", fm = "ml", main = paste("Parallel Analysis - Vignette", v))
    
    # 3. Run EFA
    efa <- fa(vignette_data, nfactors = efa_parallel$nfact, fm = "ml")
    print(efa$loadings)
    
    # 4. K-means clustering
    vignette_scaled <- scale(vignette_data)
    
    cat("Selecting optimal number of clusters (k):\n")
    nb <- fviz_nbclust(
      vignette_scaled,
      FUNcluster = kmeans,
      method = "wss",
      k.max = 6,
      nstart = 25
    )
    print(nb)
    
    optimal_k <- which.min(diff(diff(nb$data$y))) + 1
    cat("Auto-selected k =", optimal_k, "\n")
    
    set.seed(123)
    km <- kmeans(vignette_scaled, centers = optimal_k, nstart = 25)
    
    cat("K-means clustering results (k =", optimal_k, "):\n")
    print(km$size)
    
    # 5. Plot clusters
    fviz_cluster(km, data = vignette_scaled, main = paste("K-Means Clusters - Vignette", v))
    
    # === NEW: store cluster assignments ===
    cluster_assignments[[as.character(v)]] <- data.frame(
      PROLIFIC_PID = pids,
      Vignette = v,
      Cluster = km$cluster
    )
    
  } else {
    cat("Not enough complete responses for Vignette", v, "\n")
  }
}

# === NEW: Combine all vignette cluster assignments into one data frame ===
cluster_df <- bind_rows(cluster_assignments)

# Preview
head(cluster_df)

```



```{r}
# Reshape to wide format: one row per PROLIFIC_PID, columns for each Vignette
cluster_wide <- cluster_df %>%
  mutate(Vignette = paste0("V", Vignette)) %>%
  pivot_wider(
    names_from = Vignette,
    values_from = Cluster
  )

# View result
head(cluster_wide)



# Get vignette cluster columns (excluding PROLIFIC_PID)
vignette_cols <- setdiff(colnames(cluster_wide), "PROLIFIC_PID")

# All pairwise combinations of vignette cluster columns
combinations <- combn(vignette_cols, 2, simplify = FALSE)

# Run Chi-squared tests
for (pair in combinations) {
  cat("\nChi-squared test between", pair[1], "and", pair[2], "\n")
  
  tab <- table(cluster_wide[[pair[1]]], cluster_wide[[pair[2]]])
  print(tab)
  
  # Only run if the table has no 0-only rows/cols
  if (all(rowSums(tab) > 0) && all(colSums(tab) > 0)) {
    chisq <- chisq.test(tab)
    print(chisq)
  } else {
    cat("Chi-squared test not valid due to empty rows/columns.\n")
  }
}

```







# aaaa


```{r eval=FALSE}
# Convert factors to numeric if needed
# questionnaire_rating_long$PROLIFIC_PID <- as.factor(questionnaire_rating_long$PROLIFIC_PID)

# library(glmmLasso)
# 
# fit_lasso <- glmmLasso(
#   fixed = rat_mean ~ dummyPart1 + dummyPart2 + dummyPart3 + dummyPart4 + dummyPart5 + dummyPart6,
#   rnd = list(PROLIFIC_PID = ~1),
#   family = gaussian(link = "identity"),
#   data = questionnaire_rating_long,
#   lambda = 10  # tune this
# )
# 
# summary(fit_lasso)

```


```{r eval=FALSE}
library(brms)

fit_bayes <- brm(
  rat_mean ~ dummyPart1 + dummyPart2 + dummyPart3 + dummyPart4 + dummyPart5 + dummyPart6 + (1 | PROLIFIC_PID),
  data = questionnaire_rating_long,
  family = gaussian(),
  prior = set_prior("normal(0, 1)", class = "b"),
  chains = 4, iter = 2000
)

summary(fit_bayes)

```




```{r eval=FALSE}
library(relaimpo)

lm_fit <- lm(
  rat_mean ~ dummyPart1 + dummyPart3 + dummyPart5 + dummyPart6,
  data = questionnaire_rating_long[questionnaire_rating_long$Vignette %in%  c(2,3),]
)

calc.relimp(lm_fit, type = "lmg")  # "lmg" = R² decomposition

```





```{r eval=FALSE}
### save files
setwd("outputs")
#> questionnaire
## save as .xlsx file
writexl::write_xlsx(x = questionnaire, path = "questionnaire.xlsx")
## save as R object
saveRDS(questionnaire, file = "questionnaire.rds")

#> wordlistAssociations
## save as .xlsx file
writexl::write_xlsx(x = wordlistAssociations, path = "wordlistAssociations.xlsx")
## save as R object
saveRDS(wordlistAssociations, file = "wordlistAssociations.rds")

#> edgelistAssociations
## save as .xlsx file
writexl::write_xlsx(x = edgelistAssociations, path = "edgelistAssociations.xlsx")
## save as R object
saveRDS(edgelistAssociations, file = "edgelistAssociations.rds")
```

# show data

```{r eval=FALSE}
DT::datatable(questionnaire, options = list(pageLength = 5))


summary(as.numeric(table(edgelistAssociations$ID)))
table(edgelistAssociations$ID)
DT::datatable(wordlistAssociations, options = list(pageLength = 5))



DT::datatable(edgelistAssociations, options = list(pageLength = 5))
```

